Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-06-26 optim.qmd
squintability

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, $A_t$, where $t$ represents time in the sequence. The
term "geodesic" refers to maintaining the orthonormality constraint so
that each view shown is correctly a projection of the data. The PP
guided tour operates by geodesically interpolating to target planes
(projections) which have high PP index values, as provided by the
optimiser. The geodesic interpolation means that the viewer sees a
continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimised. There are
five optimisation methods implemented in the `tourr` package:

@@

From the literature, it is commonly understood that a large squint angle
implies that the objective function value is close to optimal even when
we are not very close to the perfect view to see the structure. A small
squint angle means that index function value improves substantially only
when we are very close to the perfect view. As such, low squintability
implies rapid improvement in the index value when near the perfect view.

>>>>

A tour on high-dimensional data is constructed by geodesically interpolating between pairs of planes. Any plane is described by an orthonormal basis, $A_t$, where $t$ represents time in the sequence. The term "geodesic" refers to maintaining the orthonormality constraint so that each view shown is correctly a projection of the data. The PP guided tour operates by geodesically interpolating to target planes (projections) which have high PP index values, as provided by the optimiser. The geodesic interpolation means that the viewer sees a continuous sequence of projections of the data, so they can watch patterns of interest forming as the function is optimised. There are five (unsophisticated) optimisation methods implemented in the `tourr` package:

@@

Based on the use in existing literature, it is commonly understood that a large squint angle implies that the objective function value is close to optimal even when
we are not very close to the perfect view to see the structure. A small squint angle means that index function value improves substantially only when we are very close to the perfect view. As such, low squintability implies rapid improvement in the index value when near the perfect view. For PP, a small squint angle is considered to be undesirable because it means that the optimiser needs to be very close to be able to "see" the optima. Thus, it could be difficult for the optimiser to find the optima.

### 2024-06-26 optim.tex
squintability

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, \(A_t\), where \(t\) represents time in the sequence.
The term ``geodesic'' refers to maintaining the orthonormality
constraint so that each view shown is correctly a projection of the
data. The PP guided tour operates by geodesically interpolating to
target planes (projections) which have high PP index values, as provided
by the optimiser. The geodesic interpolation means that the viewer sees
a continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimised. There are
five optimisation methods implemented in the \texttt{tourr} package:

@@

\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, coltitle=black, colback=white, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toptitle=1mm, bottomtitle=1mm, titlerule=0mm, left=2mm]

@@

From the literature, it is commonly understood that a large squint angle
implies that the objective function value is close to optimal even when
we are not very close to the perfect view to see the structure. A small
squint angle means that index function value improves substantially only
when we are very close to the perfect view. As such, low squintability
implies rapid improvement in the index value when near the perfect view.

>>>>

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, \(A_t\), where \(t\) represents time in the sequence.
The term ``geodesic'' refers to maintaining the orthonormality
constraint so that each view shown is correctly a projection of the
data. The PP guided tour operates by geodesically interpolating to
target planes (projections) which have high PP index values, as provided
by the optimiser. The geodesic interpolation means that the viewer sees
a continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimised. There are
five (unsophisticated) optimisation methods implemented in the
\texttt{tourr} package:

@@

\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, toptitle=1mm, bottomrule=.15mm, breakable, colback=white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, titlerule=0mm, left=2mm, opacityback=0, rightrule=.15mm]

@@

Based on the use in existing literature, it is commonly understood that
a large squint angle implies that the objective function value is close
to optimal even when we are not very close to the perfect view to see
the structure. A small squint angle means that index function value
improves substantially only when we are very close to the perfect view.
As such, low squintability implies rapid improvement in the index value
when near the perfect view. For PP, a small squint angle is considered
to be undesirable because it means that the optimiser needs to be very
close to be able to ``see'' the optima. Thus, it could be difficult for
the optimiser to find the optima.
